Distrivia: A Distributed Trivia Game

### 2011-05-10 papers/phase-three/paper.tex
Paper: Changes to Design and Implementation of the paper. Also finished up the abstract to give more depth. Design changes are mostly about clients and Implementation is mostly just setting up the subsections

%\title{ Distributed Systems: Project Description }
\title{ Distrivia: A Distributed Trivia Game }
\numberofauthors{3}
\author{
\alignauthor
Brian Gianforcaro \\
       \affaddr{Rochester Institute of Technology}\\
       \email{bjg1955@rit.edu}
\alignauthor
Steven Glazer \\
       \affaddr{Rochester Institute of Technology}\\
       \email{sfg6126@rit.edu}
\alignauthor
Samuel Milton \\
       \affaddr{Rochester Institute of Technology}\\
       \email{srm2997@rit.edu}
}
\maketitle

\begin{abstract}
In this paper we will discuss our decisions and detail our discoveries throughout designing a distributed trivia system.
\end{abstract}

\section{Project Motivation}
Our motivation for this project was to build a system that would allow people to compete across multiple platforms reliably.
High availability was a major concern for us, as it enables people to compete at all times.
This form of trivia is a multiplayer social game.
It takes advantage of a number of different platforms so that users can feel free to play on any device they prefer.
The web client ensures that almost any device will be able to connect and play.
Designing it on mobile platforms allows us to take advtange of the specific platform and lets people easily play on the go.
\section{Architecture}
\subsection{Servers}
For our servers, we are using Amazon EC2 micro instances.
These are running x86-64 Amazon Linux.
They have 8gb of network attached storage and 613mb of memory.

@@

Currently, we are using the NGINX \cite{nginx} in a reverse proxy configuration for our load sharing needs.
For this project there are two load balancers which are running on personal desktop computers that have reliable internet connections.
These are set up with round robin DNS entries, courtesy of ZoneEdit.com, a free DNS management service.
\section{Design}
\begin{figure}[h!]
  \centering
    \includegraphics[scale=0.4]{diagram.png}
   \caption{High level overview of the system}
 \end{figure}
\subsection{Servers}
For this project we deployed six servers running on Amazon's Elastic Compute Cloud \cite{aec}.
One server is being used as a web-monitor host running Monit\cite{monit}.
We were able to login to the web-monitor to see the status of all our servers, databases, and load sharers.
This assisted us in managing whether or not the services were running and allowed us to track down any problems, if they arose.
Each server contains the server application that responds to clients as well as the Riak \cite{riak} databases.
Riak handles auto-synchronization of the databases between the five servers.
Note that the hot spares were setup to be part of the database cluster so they have a full copy of the database on hand, however they did not accept any reads or writes directly.
This allows for them to immediately be ready for read/writes in the event that the original servers are down.
With this configuration if we were to add a server, we would just need to update Riak and it would sync the databases to any additional servers as well.
These servers, therefore, all contain the same information.
If any servers go down, the distributed system would still run fine.
We could have all servers go down except one and still be reliably handling requests.

@@

\subsection{Clients}
Our focus for clients has been on a web-based client and an Android application.
Both clients have most of the workings completed. The web-based client will
be run from desktop platforms and laptop platforms. The Android client will run
on any Android phone running OS 2.1 or higher and does not use any unique services that require specific
hardware. Both clients are nearly fully-functional. They can, currently, login/register
players, view leaderboards, join public games, and compete in public games. We are actively adding features to both clients, daily.
We have also started development on an iPhone application for distrivia. This client does not yet communicate with the server, but 
it does implement nearly all the views seen in the other clients. We have been setting up the network communication for the iPhone
platform as well.

@@

\section{Design}

\section{Implementation}

>>>>

\title{ Distrivia: A Distributed Trivia Game }
\numberofauthors{3}
\author{
\alignauthor
Brian Gianforcaro \\
       \affaddr{Rochester Institute of Technology}\\
       \email{bjg1955@rit.edu}
\alignauthor
Steven Glazer \\
       \affaddr{Rochester Institute of Technology}\\
       \email{sfg6126@rit.edu}
\alignauthor
Samuel Milton \\
       \affaddr{Rochester Institute of Technology}\\
       \email{srm2997@rit.edu}
}
\maketitle

\begin{abstract}
In this paper we will discuss our decisions and detail our discoveries throughout designing a distributed trivia system, Distrivia. 
This paper will outline our motivations behind the project, the architecture, design, implementation, lessons learned, and future work for the distributed trivia system. 
The distrivia system was implemented on 3 platforms (web, Android, and iPhone) to provide multiple ways for players to stay connected. 
It also utilizes advanced distributed systems algorithms and technology such as Amazon's EC2, Riak, SSL communication, round robin DNS, and NGINX load sharing. 
All of these services provide for high availability and secure play.
\end{abstract}

\section{Project Motivation}
Our motivation for this project was to build a system that would allow people to compete across multiple platforms reliably.
High availability was a major concern for us, as it enables people to compete at all times.
This form of trivia is a multiplayer social game.
It takes advantage of a number of different platforms so that users can feel free to play on any device they prefer.
The web client ensures that almost any device will be able to connect and play.
Designing it on mobile platforms allows us to take advtange of the specific platform and lets people easily play on the go.

\section{Architecture}
\subsection{Servers}
For our servers, we are using Amazon EC2 micro instances.
These are running x86-64 Amazon Linux.
They have 8gb of network attached storage and 613mb of memory.

@@

Currently, we are using the NGINX \cite{nginx} in a reverse proxy configuration for our load sharing needs.
For this project there are two load balancers which are running on personal desktop computers that have reliable internet connections.
These are set up with round robin DNS entries, courtesy of ZoneEdit.com, a free DNS management service.

\section{Design}
\begin{figure}[h!]
  \centering
    \includegraphics[scale=0.4]{diagram.png}
   \caption{High level overview of the system}
 \end{figure}
\subsection{Servers}
For this project we deployed six servers running on Amazon's Elastic Compute Cloud \cite{aec}.
One server is being used as a web-monitor host running Monit\cite{monit}.
We were able to login to the web-monitor to see the status of all our servers, databases, and load sharers.
This assisted us in managing whether or not the services were running and allowed us to track down any problems, if they arose.
Each server contains the server application that responds to clients as well as the Riak \cite{riak} databases.
Riak handles auto-synchronization of the databases between the five servers.
Note that the hot spares were setup to be part of the database cluster so they have a full copy of the database on hand, however they did not accept any reads or writes directly.
This allows for them to immediately be ready for read/writes in the event that the original servers are down.
With this configuration if we were to add a server, we would just need to update Riak and it would sync the databases to any additional servers as well.
These servers, therefore, all contain the same information.
If any servers go down, the distributed system would still run fine.
We could have all servers go down except one and still be reliably handling requests.

@@

\subsection{Clients}
Distrivia is implemented on three distinct platforms (web, Android, and iPhone).
The clients were designed to run without specific equipment or relying on external applications. 
The clients can be run, once installed, from desktop computers, laptops, Androids, and iPhones out of box.
The web client can be run from any browser that supports CSS, which is true of all modern web browsers.
The web client can be run from any operating system that supports web browsing and has a connection to the internet (this includes mobile devices). 
This gives users of blackberry phones and palms the ability to play, even though there is not a native app for them.
The Android client will run on any Android platform running OS 2.1 or higher and does not use any unique services that require specific hardware.
Likewise, the iPhone client will run on any iPhone/iPod touch platform running any iOS without the need for specific hardware.
The Android and iPhone clients just require an active data connection either by cell signal or wireless internet.

All of our clients were designed similarly.
We tried to keep the layout as simple as possible, for the user to be able to easily navigate the game.
The mobile clients are the most similar due to their limited screen real estate.
This was important in allowing players to play on multiple platforms while using a layout they were used to.
As mobile gaming continues to grow, our approach was to promote the play of Distrivia on the go. 
Each round can be quickly played within the course of a few minutes, which makes it ideal for playing on mobile devices.
Distrivia allows users to connect with players on any other device which gives friends the opportunity to play together without forcing them to conform to one platform.

@@

\section{Implementation}
\subsection{Network}

\subsection{Clients}
% Show images of implemented clients
Creating native apps for the Android OS and iPhone gives the users the best performance on those devices along with the native look and feel of the device for optimal design.
\subsubsection{Web}

\subsubsection{Android}

\subsubsection{iPhone}
